### 2022-08-15 main/chapter_2ljets.tex
respond to Dan's awesome comments

My main messages that this thesis reports are my contributions to the recently
published \atlas\ searches ``for new phenomena in events with two leptons,
jets, and missing transverse momentum''~\cite{atlas2022searches},
particularly to its electroweak part.
For brevity, I will call this publication $\twoljets$ and its electroweak
part $\twoljets$-electroweak.

@@

The $\twoljets$-electroweak search uses two signal models,
which we name C1N1 and GMSB and illustrate with diagrams in
Figure~\ref{fig:2ljets_signal_diagrams}.
One is an MSSM chargino-neutralino model (C1N2) that acts to produce
$\chargino_1\textrm{--}\neutralino_2$ pairs (via off-shell $W^\pm$ resonances),
that decay via $W\rightarrow q\bar q$ and $Z\rightarrow \ell^\pm \ell^\mp$
respectively to produce two leptons, jets, and two $\neutralino_1$ LSPs.
The other is a GMSB model which, from a variety of initial states, produces
$\neutralino_1\textrm{--}\neutralino_1$ pairs that decay via Higgs and $Z$
bosons, in variable branching fractions,
to final states again containing $\ell^\pm \ell^\mp$, hadronic partons,
and invisible gravitinos.
Example effects of these signal models are displayed in signal regions of the
$\twoljets$-electroweak search are displayed in
Figure~\ref{fig:2ljets_signal_examples}.
All of these regions require two leptons, jets, and
missing transverse momentum, and the signal expectations would be added on top
of the backgrounds if Nature realized these supersymmetric models.

@@

Both signal models have two free parameters that our search varies.
At each parameter point of a signal model, the search proceed by evaluating
event rates (yields) from that signal model with those parameters,
and considering what would happen of those event yields were added to a
Standard Model background.
Each signal-plus-background model is then tested against observed data in the
$\cls$ prescription.
From these tests, contours showing which parameters of those signal models are
labelled as excluded are derived, and displayed in
Figure~\ref{fig:2ljets_contours_c1n2} for C1N2 and
in~\ref{fig:2ljets_contours_gmsb} for GMSB.
In both of these figures, space to the lower left of the contours is
labelled as excluded, and space on the other side of the observed limit contour
is not excluded.
The observed contours are somewhat larger than expected due to observed
deficits in data that we have shown in Figure~\ref{fig:2ljets_summary}.

@@

\FloatBarrier
\subsection{Contributions}
Many people contributed directly and indirectly to the $\twoljets$ searches.
I claim personal responsibility for the following parts of the $\twoljets$
search and related \atlas\ efforts.
\begin{itemize}
\item Design, implementation, and execution of the electroweak part of the
analysis.
\item Service as Analysis Contact (a management role) for the $\twoljets$
project from January to October 2020.
\item Production of main data inputs for all three $\twoljets$ searches
and the RJR $3\ell$ search~\cite{atlas_rjr_3l_SUSY_2019_09},
Production of all experimental systematic variations for the background
samples;
other group members assisted with the central background sample.
Production of all electroweak signal samples and their systematic variations.
\item Evaluation of $Z$+jets uncertainties for RJR signal regions through
from simulations. These exceeded $100\%$ relative variation and supported the
use of data-driven methods.
\item Various collaborations with all other group members on common core tasks
that are shared between the searches of the $\twoljets$ project.
\item Integration of the electroweak analysis with \atlas\ combinations and
pMSSM scan efforts, by performing their validation studies and serializing the
analysis results.
\item Preparation of electroweak results for publication in the
paper~\cite{atlas2022searches} and the public HEPData
archive~\cite{maguire2017hepdata}.
\end{itemize}
Except where specified, I produced all figures presented in this thesis.
Many plotting scripts, however, are modified from versions inherited from
many past and present \atlas\ members, to whom I am grateful.

@@

The $\twoljets$-strong analysis is
detailed in Section~\ref{sec:2ljets_origins_strong} and
searches for effects from supersymmetric models that are active in the strong
sector.
It uses conventional (non-RJR) event variables and focuses on features in the
distribution of the invariant masses of lepton pairs that can have distinct
features that depend on the masses of decaying supersymmetric resonances.

@@

\FloatBarrier
\subsection{RJR}
\label{sec:2ljets_origins_rjr}

@@

The Recursive Jigsaw Reconstruction (RJR) method interprets reconstructed
particle data by reconciling them with user-specified decay trees, and operates
by making decisions of how to assign four-momenta to different `systems',
which are represented as nodes on the chosen decay graph.
Various RJR event variables are then defined in those systems' rest
frames~\cite{jackson2017sparticles, jackson2017rjr}.

Over the last several years, RJR event variables have been demonstrated to be
effective for interpreting supersymmetric particle physics
data~\cite{santoni2018probing},
and have been used in numerous \atlas\ searches~\cite{%
atlas_rjr_SUSY_2016_07,
atlas_rjr_SUSY_2016_15,
atlas_rjr_SUSY_2016_16,
atlas_rjr_23l_SUSY_2017_03,
atlas_rjr_SUSY_2018_12,
atlas_rjr_EXOT_2019_19,
atlas_rjr_3l_SUSY_2019_09
}.
They have also influenced the designs of other analyses that do not themselves
use RJR variables~\cite{atlas_rjr_mimic_SUSY_2018_06}.

@@

\FloatBarrier
\subsection{Strong}
\label{sec:2ljets_origins_strong}
Strongly supersymmetric particles, the super-partners of quarks and gluons, would
be produced at much larger cross-sections than their electroweak cousins at the
LHC.
Squarks and gluinos in simplified models are therefore excluded with masses up
to several $\eV[T]$ using hadronic data alone~\cite{atlas_susy_strong_0l}.

@@

Previous incarnations of this search have been conducted with similar
selections and target models on in older data, including partial Run~2~\cite{
atlas_susy_strong_2l_partial_run2_1,
atlas_susy_strong_2l_partial_run2,
cms_susy_2016_strong_2l_run2_1,
cms_susy_2016_strong_2l_run2
} and Run~1~\cite{
atlas_susy_strong_2l_run1,
cms_susy_2016_strong_2l_run1
} in \atlas\ and \cms.
From these Run~1 searches, one can cherry-pick excesses
of ``$3.0$ standard deviations'' in \atlas\ \cite{atlas_susy_strong_2l_run1}
and ``$2.6$ standard deviations'' in \cms~\cite{cms_susy_2016_strong_2l_run1}.
These excesses of data above estimated backgrounds were not reproduced in their
early Run~2 updates.

To target its signal models, the $\twoljets$-strong search defines seven
overlapping signal regions.
Three target on-shell $Z^{(*)}$ models by selecting $\mll$ close to $m(Z)$,
and the others are binned in $\mll$ for sensitivity to the various
kinematic endpoints that arise across the grid of various signal model
parameters.

The largest Standard Model background to this search is dileptonic $\ttbar$,
partly because no $b$-tagging veto is applied.
All three signal models produce same-flavour lepton pairs, but $\ttbar$
produces same-flavour ($ee$/$\mu\mu$) and different-flavour ($e\mu$) pairs at
equal rates.
The $\twoljets$ search exploits this difference in flavour symmetry to derive
a data-driven background estimate for all of its flavour-symmetric backgrounds,
and have its previous iterations~\cite{cms_susy_2016_strong_2l_run2_1}.
The background estimate works by taking different-flavour data, weighting them
by factors to calibrate for differences in the leptons' reconstruction, and
histogramming them in same-flavour regions.
Simulations for $\ttbar$ and the smaller flavour symmetric backgrounds
$tW$, $WW$, and $\tau\tau$ are replaced by this estimate.

@@

\FloatBarrier
\subsection{Electroweak}
\label{sec:2ljets_origins_electroweak}
As motivated in the previous sections, I have personal preferences for
simulation-driven backgrounds and orthogonal multi-bin regions.
Simulation-driven estimates are interpretable as Standard Model predictions,
relying less on loosely justified extrapolations,
and orthogonal regions have independent likelihood functions that can be
cleanly combined, by multiplication, to leverage the joint information in their
data.

@@

Additionally, $\twoljets$ widens its scope to target a GMSB simplified model
that is illustrated in Figure~\ref{fig:2ljets_signal_diagrams}.
The C1N2 and GMSB models behave similarly since both feature gaugino decays via
massive bosons ($W\kern-0.4exZ$, $ZZ$ or $Zh$) to leptons and jets, in
similarly structures decay trees to invisible supersymmetric pairs.
Although widening sensitivity to both models must sacrifice some sensitivity to
each individually, the increased degree of model independence is itself of
value for future reinterpretations of our data.
We are happy to sacrifice some precision for generality.
Since the $W$ and $Z$ boson masses are similar, this sacrifice turns out to be
quite small anyway.

Multi-bin fits are not special.
They are standard \atlas\ practice, and there should have been no complication
had I designed copied SR2-int and SR2-high for $\twoljets$-electroweak but with
with minor tweaks to change them into orthogonally binned regions.
However, my design decisions for $\twoljets$-electroweak became much more
ambitious, increasing to a total of $13$ signal regions with complicated
definitions.
Although this multi-bin design squeezed out some extra sensitivity,
it also caused problems --- for the clarity of communication and the accuracy
of our background modelling;
these issues shall be developed throughout the remainder of this chapter.

@@

Unfortunately, the $\twoljets$-electroweak regions designs were finalized
before unblinding and before I appreciated this problem with background
uncertainties in the smaller regions.
Various workaround are invoked to assign plausible variations despite
small yields and failing statistical precision, which shall be seen in the
remainder of this chapter.
As seen in the splash summary of Section~\ref{sec:2ljets_splash}, the
$\twoljets$-electroweak search observes agreement with post-fit backgrounds
except for some local deficits of data below backgrounds.
Exclusion contours exceed their expected range because of these deficits,
but they can only be trusted if the background model is trusted.

@@

The `$\twoljets$' name is intended to indicate interest in events with two
charged leptons (those are electrons or muons --- not taus), and at least one
jet.
Though not in the name, an $\met$ requirement is also implicit from the context
of $R$-parity conserving supersymmetric models.
This Section~\ref{sec:2ljets_events} serves to describe how events are selected
for the $\twoljets$ analyses, and which event variables are extracted for use
in its selections.

Our dataset comprises $136.0\,\mathrm{fb}^{-1}$ ($\pm 1.7\%$) of integrated
luminosity observed during normal operations of the \atlas\ detector in the
years $2015$ through to $2018$ of the LHC Run~2~\cite{DAPR-2018-01}.
Triggers and object definitions are shared between the three $\twoljets$
searches.
Higher level event variables and selections will only be described here for the
$\twoljets$-electroweak search.

@@

Each trigger in \atlas' arsenal activates on a selection comprising some
combination of sufficiently hard objects.
The set of targeted objects includes all those used in this analysis and more:
electron, muons, jets, $b$-jets, $\met$, as well as photons and hadronically
decaying tau leptons~\cite{atlas_PERF_2007_01}.

@@

This complicated collection of trigger requirements motivates lepton selections
that exceed these trigger thresholds,
such that they avoid the need to accurately model their jagged activation
thresholds and the variable efficiencies near to those thresholds~\cite{
atlas_trigger_egamma_run2,
atlas_trigger_muon_run2
}.
To avoid the most of these thresholds, all leptons selected for the
$\twoljets$-electroweak regions are required to have $\pt > 25\,\eV[G]$.

@@

The benefit of an additional trigger could be in increased signal yields if
that trigger does not also harmfully increase inseparable backgrounds.
A study, \emph{performed by Knut Vadla}, investigated the effects
of including one-lepton triggers.
In a signal-region-like selection requiring two leptons, two jets and
$\met > 200\,\eV[G]$, this study showed that two-lepton triggers already accept
in excess of $90\%$ of events from benchmark C1N2 models.
Therefore, any additional trigger has at most a $10\%$;
improvement. Adding one-lepton triggers was found to increase efficiency by
only a few percentage points, depending on the model.
We in the $\twoljets$ team considered that to be too little to be worthwhile.

@@

Reconstructed lepton objects feature momenta, charges, and flavours.
Each lepton is accepted for use in $\twoljets$ only if it satisfies certain
kinematic and is identified with `quality' criteria.
For both electron and muons, these criteria are standardized with labels
`loose', `medium', and `tight';
these labels are nested in the sense that tight implies medium
and medium implies loose.
Other, more specialized, criteria exist in \atlas, but they are not relevant
$\twoljets$.

@@

Leptons in $\twoljets$ are further sliced into two classes:
`baseline' and `signal'.
Baseline leptons and are used in background estimation methods and to ensure
orthogonality with other \atlas\ analyses that study events with differing
lepton counts, such that our results might be combined in future to deliver an
improved total sensitivity.
Baseline requirements are softer and lower quality than signal requirements.

@@

Jets in $\twoljets$ use the anti-$k_t$ algorithm with the standard small-radius
parameter of $R=0.4$~\cite{jet_anti_kt}.
Jets are formed from `topological' clusters of energy deposits in the
calorimeters~\cite{atlas_jet_topo_PERF_2014_07}.

@@

Once we know about all the reconstructed leptons and hadronic jets, we can
also improve our rejection of non-prompt leptons from hadron decays, since
leptons near to jets become suspect.
To remove duplicates and further reduce non-prompt lepton contributions, we
apply a standard overlap-removal procedure.
This procedure, which is detailed in the $\twoljets$
paper~\cite{atlas2022searches},
comprises a sequence of rules that modify the collection of event objects.
Some rules decide which object to keep from an overlapping pair,
and others remove leptons overlapping with jets based on their $\Delta R$
and $\pt$.

@@

All signal models in $\twoljets$ generate hard, invisible particles.
By selecting for large $\met$, however, we select events in which either there
were real invisible particles or something was badly mismeasured.
And the rate of bad mismeasurements from Standard Model processes
with large cross-sections can be substantial.
A standard selection requirement in $\twoljets$-electroweak is
$\met > 100\,\eV[G]$.
This is a nice round number (in these conventional units), and corresponds
roughly to where the $\met$-less background $\zjets$ tails off,
and also corresponds roughly to several standard deviations of
a typical $\met$ noise scale of approximately $10\textrm{--}30\,\eV[G]$;
I observe that noise scale by inspecting $\met / \metsig$ in
$\twoljets$-electroweak selections, where $\metsig$ is defined in the next
section.

@@

However, alternatives did select different subsets of signal yields:
the alternative jet assignments had some potential to define secondary signal
regions to catch those signal events which fall outside the primary
$\mjj$ window.
One can, for example, select $\mjj \notin 60\textrm{--}110\,\eV[G]$ and
$\mjj^\mathrm{alternative} \in 60\textrm{--}110\,\eV[G]$ for some additional
sensitivity.
However, this comes at the cost of complexity, particularly in the boundaries
with control and validation regions, and is therefore not used here.

@@

Furthermore, simplicity encourages interpretability and trust.
A selection at $100\,\eV[G]$ may be slightly outperformed in precision by one
at $100.314\,\eV[G]$, but the former is preferable --- it is easily
communicated and perhaps understood or reproduced by readers.
The latter could be overly tuned, perhaps by a learning algorithm, to specific
simulations; perhaps it cut out a highly weighted sample that, although
annoying, was an
honest representation of our modelling.
Restricting ourselves to round numbers (in the units with which we communicate)
not only clarifies our communications,
but keeps us honest by denying ourselves the ability to over-fit simulations.

@@

\subsection{High}
\label{sec:2ljets_high}
The highest $\metsig$ selections, starting around $\metsig > 18$, contain
pure \diboson\ backgrounds in relatively small quantities, along with the bulk
of signal samples at the heaviest resonance masses.
This is therefore the most promising territory for signal regions that are
highly sensitive to the most extreme signals.
Their backgrounds, however, must be constrained by control regions in
less-extreme parameter spaces with sufficiently many data to precisely
normalize their scale factors.

Diboson backgrounds and C1N2 signals have more in common than just $\met$;
both also produce resonant $Z\rightarrow \ell^+\ell^-$ pairs
and do not enhance the rate of heavy-flavour jets, as displayed in
Figure~\ref{fig:2ljets_high_mll_b}.
Since signals and backgrounds have such similar shapes in $\mll$ and $\nbtag$,
sensitivity is only decreased by rejecting portions of these distributions.
Although most $\twoljets$-electroweak regions use tighter selection on these
variables to reduce backgrounds, this observation allows us to win
additional sensitivity in High regions by loose requirements
of $\mll \in 71\textrm{--}111$ and $\nbtag \leq 1$.
This $\mll$ window loosely selects the $m(Z)$ peak to scoop up nearby rare or
mismeasured events away from the central peak.
Similarly, allowing up to one $b$-tag improves yields without boosting
backgrounds, since the $\metsig$ requirement already removes most $\ttbar$
processes.

@@

As explained in Section~\ref{sec:2ljets_mt2}, requiring $\mttwo$ above $m(W)$
effectively removes $\ttbar$ and other backgrounds.
The relatively low requirement $\mttwoll > 80\,\eV[G]$ is chosen here and
elsewhere in the $\twoljets$-electroweak design to reduce these backgrounds
while minimize lost signal yields.

To maximize projected sensitivity, SR-High is split into four bins with
one split at $\rjj = 0.8$, to form SR-High-8 and SR-High-16, and a second at
$\metsig = 21$ to split these into SR-High-*-a and SR-High-*-b.
For example, SR-High-8-b requires $\rjj < 0.8$ and $\metsig > 21$.
The regions SR-High-8 and SR-High-16 are displayed in
Figure~\ref{fig:2ljets_high_region}, with their -a and -b bins visible across
the $\metsig$ distribution.

@@

Total backgrounds predict data well in VR-$\llbb$.
What was not appreciated soon enough, however, was an excess in the $\metsig$
tail of this region, which is shown in the $\twoljets$ paper and reproduced
Figure~\ref{fig:2ljets_high_metsig_vrllbb_paper}.
The highest bins have very small background predictions and positive data.
Had this been noticed before unblinding, more validation studies or altered
modelling might have been necessary.
``In the end'', commented one \atlas\ author~\cite{comment2022vrllbb},
``it seems fine since there's not events in the SR!''

@@

\FloatBarrier
\subsection{Intermediate}
\label{sec:2ljets_int}

>>>>

The main messages that this thesis reports are my contributions to the recently
published \atlas\ searches ``for new phenomena in events with two leptons,
jets, and missing transverse momentum''~\cite{atlas2022searches},
particularly to its electroweak part.
For brevity, I will call this publication $\twoljets$ and its electroweak
part $\twoljets$-electroweak.

@@

The $\twoljets$-electroweak search uses two signal models,
which we name C1N1 and GMSB and illustrate with diagrams in
Figure~\ref{fig:2ljets_signal_diagrams}.
One is an MSSM chargino-neutralino model (C1N2) that acts to produce
$\chargino_1\textrm{--}\neutralino_2$ pairs (via off-shell $W^\pm$ resonances),
that decay via $W\rightarrow q\bar q$ and $Z\rightarrow \ell^\pm \ell^\mp$
respectively to produce two leptons, jets, and two $\neutralino_1$ LSPs.
The other is a GMSB model which, from a variety of initial states, produces
$\neutralino_1\textrm{--}\neutralino_1$ pairs that decay via Higgs and $Z$
bosons, with variable branching fractions,
to final states again containing $\ell^\pm \ell^\mp$, hadronic partons,
and invisible gravitinos.
Example effects of these signal models are displayed in signal regions of the
$\twoljets$-electroweak search are displayed in
Figure~\ref{fig:2ljets_signal_examples}.
All of these regions require two leptons, jets, and
missing transverse momentum, and the signal expectations would be added on top
of the backgrounds if Nature realized these supersymmetric models.

@@

Both signal models have two free parameters that our search varies.
At each parameter point of a signal model, the search proceed by evaluating
event rates (yields) from that signal model with those parameters,
and considering what would happen of those event yields were added to a
Standard Model background.
Each signal-plus-background model is then tested against observed data in the
$\cls$ prescription.
From these tests, contours showing which parameters of those signal models are
labelled as excluded are derived, and displayed in
Figure~\ref{fig:2ljets_contours_c1n2} for C1N2 and in
Figure~\ref{fig:2ljets_contours_gmsb} for GMSB.
In both of these figures, space to the lower left of the contours is
labelled as excluded, and space on the other side of the observed limit contour
is not excluded.
The observed contours are somewhat larger than expected due to observed
deficits in data that we have shown in Figure~\ref{fig:2ljets_summary}.

@@

\subsection{Contributions}
Many people contributed directly and indirectly to the $\twoljets$ searches.
I claim personal responsibility for the following parts of the $\twoljets$
search and related \atlas\ efforts.
\begin{itemize}
\item Design, implementation, and execution of the electroweak part of the
analysis.
\item Service as Analysis Contact (a management role) for the $\twoljets$
project from January to October 2020.
\item Production of main data inputs for all three $\twoljets$ searches
and the RJR $3\ell$ search~\cite{atlas_rjr_3l_SUSY_2019_09}.
Production of all experimental systematic variations for the background
samples --- other group members assisted with the central background sample.
Production of all electroweak signal samples and their systematic variations.
\item Evaluation of $Z$+jets uncertainties for RJR signal regions from
simulations. Their variations exceeded $100\%$ relative variation and supported
the use of data-driven methods.
\item Various collaborations with all other group members on common core tasks
that are shared between the searches of the $\twoljets$ project.
\item Integration of the electroweak analysis with \atlas\ combinations and
pMSSM scan efforts, by performing their validation studies and serializing the
analysis results.
\item Preparation of electroweak results for publication in the
paper~\cite{atlas2022searches} and the public HEPData
archive~\cite{maguire2017hepdata}.
\end{itemize}
Except where specified, I produced all figures presented in this thesis.
Many plotting scripts, however, are modified from versions inherited from
many past and present \atlas\ members, to whom I am grateful.

@@

The $\twoljets$-strong analysis is
detailed in Section~\ref{sec:2ljets_origins_strong} and
searches for effects from supersymmetric models that are active in the strong
sector.
It uses conventional (non-RJR) event variables and focuses on features in the
distribution of the invariant masses of lepton pairs.
That dilepton mass distribution can have distinct features that would stand
out from Standard Model backgrounds and depend on the masses of decaying
supersymmetric resonances.

@@

\subsection{RJR}
\label{sec:2ljets_origins_rjr}
The Recursive Jigsaw Reconstruction (RJR) method interprets reconstructed
particle data by reconciling them with user-specified decay trees, and operates
by making decisions of how to assign four-momenta to different `systems',
which are represented as nodes on the chosen decay graph.
Various RJR event variables are then defined in those systems' rest
frames~\cite{jackson2017sparticles, jackson2017rjr}.

Over the last several years, RJR event variables have been demonstrated to be
effective for interpreting supersymmetric particle physics
data~\cite{santoni2018probing},
and have been used in numerous \atlas\ searches~\cite{
atlas_rjr_SUSY_2016_07,
atlas_rjr_SUSY_2016_15,
atlas_rjr_SUSY_2016_16,
atlas_rjr_23l_SUSY_2017_03,
atlas_rjr_SUSY_2018_12,
atlas_rjr_EXOT_2019_19,
atlas_rjr_3l_SUSY_2019_09
}.
They have also influenced the designs of other analyses that do not themselves
use RJR variables~\cite{atlas_rjr_mimic_SUSY_2018_06}.

@@

@@

\FloatBarrier
\subsection{Strong}
\label{sec:2ljets_origins_strong}
Strongly interacting supersymmetric particles, the super-partners of quarks and
gluons, would be produced at much larger cross-sections than their electroweak
cousins at the LHC.
Squarks and gluinos in simplified models are therefore excluded with masses up
to several $\eV[T]$ using hadronic data alone~\cite{atlas_susy_strong_0l}.

@@

Previous incarnations of this search have been conducted with similar
selections and signal models in older data, including partial Run~2~\cite{
atlas_susy_strong_2l_partial_run2_1,
atlas_susy_strong_2l_partial_run2,
cms_susy_2016_strong_2l_run2_1,
cms_susy_2016_strong_2l_run2
} and Run~1~\cite{
atlas_susy_strong_2l_run1,
cms_susy_2016_strong_2l_run1
} in \atlas\ and \cms.
From these Run~1 searches, one can cherry-pick excesses
of ``$3.0$ standard deviations'' in \atlas\ \cite{atlas_susy_strong_2l_run1}
and ``$2.6$ standard deviations'' in \cms~\cite{cms_susy_2016_strong_2l_run1}.
These excesses of data above estimated backgrounds were not reproduced in their
early Run~2 updates.

To target its signal models, the $\twoljets$-strong search defines seven
overlapping signal regions.
Three target on-shell $Z^{(*)}$ models by selecting $\mll$ close to
$m(Z)\approx 91.2\,\eV[G]$~\cite{pdg2020review},
and the others are binned in $\mll$ for sensitivity to the various
kinematic endpoints that arise across the grid of various signal model
parameters.

The largest Standard Model background to this search is dileptonic $\ttbar$,
partly because it employes no veto for $b$-tagged jets.
All three signal models produce same-flavour lepton pairs, but $\ttbar$
produces same-flavour ($ee$/$\mu\mu$) and different-flavour ($e\mu$) pairs at
equal rates.
The $\twoljets$ search exploits this difference in flavour symmetry to derive
a data-driven background estimate for all of its flavour-symmetric backgrounds,
and have its previous iterations~\cite{cms_susy_2016_strong_2l_run2_1}.
The background estimate works by taking different-flavour data, weighting them
by factors to calibrate for differences in the leptons' reconstruction, and
histogramming them in same-flavour regions.
Simulations for $\ttbar$ and the smaller flavour symmetric backgrounds
$tW$, $WW$, and $\tau\tau$ are replaced by this estimate.

@@

\subsection{Electroweak}
\label{sec:2ljets_origins_electroweak}
As motivated in the previous sections, I have personal preferences for
simulation-driven backgrounds and orthogonal multi-bin regions.
Simulation-driven estimates are interpretable as Standard Model predictions,
relying less on loosely justified extrapolations,
and orthogonal regions have independent likelihood functions that can be
cleanly combined, by multiplication, to leverage the joint information in their
data.

@@

Additionally, $\twoljets$ widens its scope to target a GMSB simplified model
that is illustrated in Figure~\ref{fig:2ljets_signal_diagrams}.
The C1N2 and GMSB models behave similarly since both feature gaugino decays via
massive bosons ($W\kern-0.4exZ$, $ZZ$ or $Zh$) to leptons and jets, in
similarly structures decay trees to invisible supersymmetric pairs.
Although widening sensitivity to both models must sacrifice some sensitivity to
each individually, the increased degree of model independence is itself of
value for future reinterpretations of our data.
We are happy to sacrifice some precision for generality.
Since the $W$ and $Z$ boson masses are similar, this sacrifice turned out to
have quite a minor impact.

Multi-bin fits are not special.
They are standard \atlas\ practice, and there should have been no complication
had I imitated SR2-int and SR2-high for $\twoljets$-electroweak but with minor
tweaks to change them into orthogonally binned regions.
However, my design decisions for $\twoljets$-electroweak became much more
ambitious, increasing to a total of $13$ signal regions with complicated
definitions.
Although this multi-bin design squeezed out some extra sensitivity,
it also caused problems --- for the clarity of communication and the accuracy
of our background modelling;
these issues shall be developed throughout the remainder of this chapter.

@@

Unfortunately, the $\twoljets$-electroweak regions designs were finalized
before unblinding, and before I appreciated this problem with background
uncertainties in the smaller regions.
Various workarounds are therefore invoked to assign plausible variations
despite small yields and failing statistical precision, which shall be seen in
the remainder of this chapter.
As seen in the splash summary of Section~\ref{sec:2ljets_splash}, the
$\twoljets$-electroweak search observes agreement with post-fit backgrounds
except for some local deficits of data below backgrounds.
Exclusion contours exceed their expected range because of these deficits,
but they can only be trusted if the background model is trusted.

@@

The `$\twoljets$' name is intended to indicate interest in events with two
charged leptons (those are electrons or muons --- not taus), and at least one
jet.
Though not in the name, an $\met$ requirement is also implicit from the context
of $R$-parity conserving supersymmetric models.
This Section~\ref{sec:2ljets_events} serves to describe how events are selected
for the $\twoljets$ searches, and which event variables are extracted for use
in its selections.

Our dataset comprises $139.0\,\mathrm{fb}^{-1}$ ($\pm 1.7\%$) of integrated
luminosity observed during normal operations of the \atlas\ detector in the
years $2015$ through to $2018$ of the LHC Run~2~\cite{DAPR-2018-01}.
Triggers and object definitions are shared between the three $\twoljets$
searches.
Higher level event variables and selections will only be described here for the
$\twoljets$-electroweak search.

@@

Each trigger in \atlas' arsenal activates on a selection comprising some
combination of sufficiently hard objects.
The set of targeted objects includes all those used in this analysis and more:
electrons, muons, jets, $b$-jets, $\met$, as well as photons and hadronically
decaying tau leptons~\cite{atlas_PERF_2007_01}.

@@

This complicated collection of trigger requirements motivates lepton selections
that exceed these trigger thresholds,
such that they avoid the need to accurately model their jagged activation
thresholds and varying efficiencies near to those thresholds~\cite{
atlas_trigger_egamma_run2,
atlas_trigger_muon_run2
}.
To avoid the most of these thresholds, all leptons selected for the
$\twoljets$-electroweak regions are required to have $\pt > 25\,\eV[G]$.

@@

The benefit of an additional trigger could be in increased signal yields if
that trigger does not also harmfully increase inseparable backgrounds.
A study, \emph{performed by Knut Vadla}, investigated the effects
of including one-lepton triggers.
In a signal-region-like selection requiring two leptons, two jets and
$\met > 200\,\eV[G]$, this study showed that two-lepton triggers already accept
in excess of $90\%$ of events from benchmark C1N2 models.
Therefore, any additional trigger has at most a $10\%$ improvement.
Adding one-lepton triggers was found to increase acceptance rates by only a
few percentage points.
Especially since this must come with a comparable increase in background rates,
the benefit is too little to be considered worthwhile.

@@

Reconstructed lepton objects feature momenta, charges, and flavours.
Each lepton is accepted for use in $\twoljets$ only if it satisfies certain
kinematic selections and is identified with `quality' criteria.
For both electron and muons, these criteria are standardized with labels
`loose', `medium', and `tight';
these labels are nested in the sense that tight implies medium
and medium implies loose.
Other, more specialized, criteria exist in \atlas, but they are not relevant
$\twoljets$.

@@

Leptons in $\twoljets$ are further sliced into two classes:
`baseline' and `signal'.
Baseline leptons and are used in background estimation methods and to ensure
orthogonality with other \atlas\ analyses that study events with differing
lepton counts, such that our results can be combined in future to deliver an
improved total sensitivity.
Baseline requirements are softer and lower quality than signal requirements.

@@

Jets in $\twoljets$ are constructed with the anti-$k_t$ algorithm with a
standard small radius parameter of $R=0.4$~\cite{jet_anti_kt}.
Jets are formed from `topological' clusters of energy deposits in the
calorimeters~\cite{atlas_jet_topo_PERF_2014_07}.

@@

Once we know about all the reconstructed leptons and hadronic jets, we can
also improve our rejection of non-prompt leptons from hadron decays by
leveraging the tendancy for those non-prompt leptons to be near to jets.
To remove duplicates and further reduce non-prompt lepton contributions, we
apply a standard overlap-removal procedure.
This procedure, which is detailed in the $\twoljets$
paper~\cite{atlas2022searches},
comprises a sequence of rules that modify the collection of event objects.
Some rules decide which object to keep from an overlapping pair,
and others remove leptons overlapping with jets based on their $\Delta R$
and $\pt$.

@@

All signal models in $\twoljets$ generate hard, invisible particles.
By selecting for large $\met$, however, we select events with either
real invisible particles or bad mismeasurements.
And the rate of badly measured events from Standard Model processes with large
cross-sections can be substantial.
A standard selection requirement in $\twoljets$-electroweak is
$\met > 100\,\eV[G]$.
This is a nice round number (in these conventional units), and corresponds
roughly to where the $\met$-less background $\zjets$ tails off,
and also corresponds roughly to several standard deviations of
a typical $\met$ noise scale of approximately $10\textrm{--}30\,\eV[G]$.
(I observe that noise scale by inspecting $\met / \metsig$ in
$\twoljets$-electroweak selections, where $\metsig$ is defined in the next
section.)

@@

However, alternative jet pair assignments did select different subsets of
signal yields:
the alternative jet assignments had some potential to define secondary signal
regions to catch those signal events which fall outside the primary
$\mjj$ window.
One can, for example, select $\mjj \notin 60\textrm{--}110\,\eV[G]$ and
$\mjj^\mathrm{alternative} \in 60\textrm{--}110\,\eV[G]$ for some additional
sensitivity.
However, this comes at the cost of complexity, particularly in the boundaries
with control and validation regions, and is therefore not used here.

@@

Furthermore, simplicity encourages interpretability and trust.
A selection at $100\,\eV[G]$ may be slightly outperformed in precision by one
at $100.314\,\eV[G]$, but the former is preferable --- it is easily
communicated and perhaps understood or reproduced by readers.
The latter could be overly tuned, perhaps by a learning algorithm, to specific
simulations; perhaps it cut out a highly weighted sample that, although
annoying, was an
honest representation of our modelling.
Restricting ourselves to round numbers (in the units with which we communicate)
not only clarifies our communications,
but keeps us honest by restricting our ability to over-fit to simulations.

@@

\subsection{High}
\label{sec:2ljets_high}
Our highest $\metsig$ selections, starting around $\metsig > 18$, contain
pure \diboson\ backgrounds in relatively small quantities, along with the bulk
of signal samples at the heaviest resonance masses.
This is therefore the most promising territory for signal regions that are
highly sensitive to the most extreme signals.
Their backgrounds, however, must be constrained by control regions in
less-extreme parameter spaces with sufficiently many data to precisely
normalize their scale factors.

Diboson backgrounds and C1N2 signals have more in common than just $\met$;
both also produce resonant $Z\rightarrow \ell^+\ell^-$ pairs
and do not enhance the rate of heavy-flavour jets, as displayed in
Figure~\ref{fig:2ljets_high_mll_b}.
Since signals and backgrounds have such similar shapes in $\mll$ and $\nbtag$,
sensitivity is only decreased by rejecting portions of these distributions.
Although most $\twoljets$-electroweak regions use tighter selection on these
variables to reduce backgrounds, this observation allows us to win
additional sensitivity in High regions by loose requirements
of $\mll \in 71\textrm{--}111\,\eV[G]$ and $\nbtag \leq 1$.
This $\mll$ window loosely selects the $m(Z)$ peak to scoop up nearby rare or
mismeasured events away from the central peak.
Similarly, allowing up to one $b$-tag improves yields without boosting
backgrounds, since the $\metsig$ requirement already removes most $\ttbar$
processes.

@@

As explained in Section~\ref{sec:2ljets_mt2}, requiring $\mttwo$ above $m(W)$
effectively removes $\ttbar$ and other backgrounds.
The relatively low requirement $\mttwoll > 80\,\eV[G]$ is chosen here and
elsewhere in the $\twoljets$-electroweak design to reduce these backgrounds
while minimizing lost signal yields.

To maximize our projected sensitivity, SR-High is split into four bins with
one division at $\rjj = 0.8$, to form SR-High-8 and SR-High-16, and a second
division at $\metsig = 21$ to split each of those into
SR-High-*-a and SR-High-*-b.
For example, SR-High-8-b requires $\rjj < 0.8$ and $\metsig > 21$.
The regions SR-High-8 and SR-High-16 are displayed in
Figure~\ref{fig:2ljets_high_region}, with their -a and -b bins visible across
the $\metsig$ distribution.

@@

Overall, the total background yield predict the data well in VR-$\llbb$.
What was not appreciated soon enough, however, was an excess in the $\metsig$
tail of this region, which is shown in the $\twoljets$ paper and reproduced
Figure~\ref{fig:2ljets_high_metsig_vrllbb_paper}.
The highest bins have small background predictions and positive data.
Had this been noticed before unblinding, more validation studies or altered
modelling might have been necessary.
``In the end'', commented one \atlas\ author~\cite{comment2022vrllbb},
``it seems fine since there's not events in the SR!''

@@

\subsection{Intermediate}
\label{sec:2ljets_int}
